---
title: "mapedit"
author: "Jamie Montgomery"
date: "2/4/2020"
output: html_document
---

# Overview

This is a short introduction to using the `mapedit` R package developed by Tim Applehans and others. For more information I encourage you to check out: [https://www.r-spatial.org/r/2019/03/31/mapedit_leafpm.html](https://www.r-spatial.org/r/2019/03/31/mapedit_leafpm.html)


## Setup

You will need the `sf`, `mapedit` and `leaflet` packages.

```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(sf)
library(mapedit)
library(leaflet)
```

## Load data

The `seaweed_range_lines` shapefile contains multiple polyline objects that identify native ranges of seaweed around the world. I created it by hand for a paper on [seaweed aquaculture potential (Froehlich et al. 2019)](https://www.sciencedirect.com/science/article/abs/pii/S0960982219308863).

```{r}
seaweed <- read_sf("data/seaweed_range_lines.shp")
```

## Create a spatial object from scratch

Open up a blank leaflet map using the `editMap()` function and begin building lines, polygons or points. When done click finished and you have a list. To get the final as a simple feature select the data $finished.

```{r}

seaweed_lines <- editMap(leaflet() %>% addTiles())
sw_lines <- seaweed_lines$finished

#save using write_sf
write_sf(sw_lines, "../../Data/seaweed_range_lines", driver = "ESRI Shapefile")
```

## Edit a shapefile

Can I edit a shapefile? (answer: yes!)

Let's edit the seaweed lines shapefile
```{r}
new_lines <- mapview(seaweed) %>%
  editMap("seaweed")
```

